Optimizing Abstract$^{\text 2}$ Machines

### 2012-11-08 paper/paper.tex

\documentclass[preprint]{sigplanconf}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
% \usepackage{xltxtra}
% \setmonofont[Scale=MatchLowercase]{DejaVu Sans Mono}
%
\input{preamble}

@@

>>>>

\documentclass[preprint]{sigplanconf}
\usepackage{alltt}
\usepackage{amsmath,amsthm}
\usepackage{amssymb}
\usepackage{stmaryrd}

\newtheorem{theorem}{Theorem}
% \usepackage{xltxtra}
% \setmonofont[Scale=MatchLowercase]{DejaVu Sans Mono}
%
\input{preamble}

@@

\appendix
\section{Proof: laziness is precision-preserving}
Given widened machine configurations, we can show that lazy
non-determinism is precision-preserving in the cases that application
positions are store-allocated and not. We first show the latter since
it uses less cluttered transition rules. The high level is that the
``fan-out'' of non-determinism is collapsed back in the store (in the
store-allocated applications case) or delayed a single step by
laziness that would have happened in a step of strictness.

Details of lazy machine given as a diff from the strict machine. \\
State space of $lazy-\widehat{CESK}^*_t$ varies just in
$\mathit{Value}$ (storable values do not change):
\begin{align*}
\mval \in \mathit{Value} &::= z \mid b \mid o \mid \\saddr{\maddr} \\
\end{align*}

We define a relation between strict and lazy machines that illustrates
the quotient that the lazy states

\newcommand{\lapprx}[3]{#1 \approx_{#2} #3}
\newcommand{\apprx}[3]{#1 \sim_{#2} #3}
\newcommand{\capprx}[2]{#1 \sim #2}

\begin{align*}
\ev{\mexp, \menv, \mkont, \mcntr} \in cs &\Rightarrow \lapprx{\ev{\mexp, \menv,\mkont, \mcntr}}{\msto}{cs} \\
\co{\mkont, \mval} \in cs \mbox{ and } \mval' \in force(\msto, \mval) &\Rightarrow \lapprx{\co{\mkont, \mval}}{\msto}{cs} \\
\ap{\mval_0, \mval_1, \mkont} \in cs \mbox{ and } \mval_i' \in force(\msto, \mval_i) &\Rightarrow \lapprx{\ap{\mval_0', \mval_1', \mkont}}{\msto}{cs} \\
\ev{\mexp, \menv, \mkont} \in lcs &\Rightarrow \apprx{\ev{\mexp, \menv,\mkont}}{\msto}{lcs} \\
\co{\mkont, \mval} \in lcs \mbox { and } \mval' \in force(\msto, \mval) &\Rightarrow \apprx{\co{\mkont, \mval'}}{\msto}{lcs} \\
\ap{\mval_0, \mval_1, \mkont} \in lcs \mbox { and } \mval_i' \in force(\msto, \mval_i) &\Rightarrow \apprx{\ap{\mval_0', \mval_1', \mkont}}{\msto}{lcs} \\
\forall lc \in lcs, \lapprx{lc}{\msto}{cs}, \mbox{ and }\forall c \in cs, \apprx{c}{\msto}{lcs}. &\Rightarrow \capprx{(lcs, \msto)}{(cs, \msto)}
\end{align*}

We also need metafunctions for adding and removing a store component
from states. These are straightforwardly defined, so just name them
$wn$ and $nw$ respectively.
%% c2vc(<e^l r k d>,s) = <e^l r s k d>
%% c2vc(<v k d>,s) = <v s k d>
%% vc2c(<e^l r s k d>) = <e^l r k d>
%% vc2c(<v s k d>) = <v k d>

The property we show is that all states stay related through
reduction. The main difficulty is in showing the resulting stores are
in fact equal. We do this by showing each individual reduction takes
equal stores to equal stores, and that the union of all these
reductions is the full reduction.

\begin{theorem}
$\forall lcs,cs,lcs',\msto,\msto'$ if $\capprx{(lcs, \msto)}{(cs,\msto)}$
  and $(lcs,\msto) \machstep (lcs',\msto')$ then there exists $cs'$
  such that $(cs,\msto) \machstep (cs',\msto')$
\end{theorem}
\begin{proof}
...
\end{proof}

%% --Eval
%% <x r s k d> --> <v s k d> where v in s(r(x))  [VAR]
%% <(e0^l0 e1^l1)^{l,lf,la} r s k d> --> <e0^l0 r s\sqcup[ld |-> {k}] ar^{l,lf,la}(e1^l1 r ld) d> [APP]
%% <(lambda x e) r s k d> --> <clos(x e r) s k d> [CLOS]
%% --Continue
%% <v s ar^{l,lf,la}(e^le r a) d> --> <e^le r s\sqcup[lfd |-> {v}] fn^{l,la}(lfd, a)> [CO]
%% --Apply
%% <v s fn^{l,la}(fa, ka) d> --> <e^fl r[x |-> xd'] s'[xd' |-> s'(lad)] k d'> [AP]
%%   where clos(x e^fl r) in s(fa)
%%         k in s(ka)
%%         s' = s\sqcup[lad |-> {v}]
%%         d' = truncate(ld, K)

%% This extra binding corresponds directly to the intermediate
%% bindings introduced by ANF.
%% Matt's ANF analyses likely have all the flavours of our "laziness" technique, but this is how we get our chocolate with our peanut butter as Olin would say.

%% Compare this to lazyCESK^*t:
%% v ::= clos(x e r) | addr(a)
%% Clos ::= clos(x e r)
%% s : a -> P(Clos)
%% others the same

%% force(s, addr(a)) = s(a)
%% force(s, clos(x e r)) = {clos(x e r)}

%% --Eval
%% <x r s k d> --> <addr(r(x)) s k d>
%% others the same
%% --Continue
%% <v s ar^{l,lf,la}(e^le r a) d> --> <e^le r s\sqcup[lfd |-> force(s,v)] fn^{l,la}(lfd, a)>
%% --Apply
%% <v s fn^{l,la}(fa, ka) d> --> <e^fl r[x |-> xd'] s'[xd' |-> s'(lad)] k d'>
%%   where clos(x e^fl r) in s(fa)
%%         k in s(ka)
%%         s' = s\sqcup[lad |-> force(s, v)]
%%         d' = truncate(ld, K)
%% The same globalizing stuff applies.

%% Let us relate the state spaces of the two abstract machines for the lambda calculus in the following way

%% Then we can show that if lazyconf ~ conf and lazyconf ==> lazyconf' then there exists a conf' such that conf ==> conf' and lazyconf' ~ conf'
%% Let lc in lazyconf.cs be arbitrary (call lazyconf.s, s)
%% By cases on lc:
%% <x r k d>:
%% Premises:
%% <x r k d> in conf.cs
%% Thus:
%% Since
%% <x r s k d> -->lazy   <addr(r(x)) s k d>
%% <x r s k d> -->strict <v s k d> where v in s(r(x))
%% By definition of force, v in force(addr(r(x)), s)
%% Thus the first rule of ~_s applies and the second rule of l~_s applies.

%% <(lambda x e) r k d>:
%% Trivial

%% <(e0^l0 e1^l1)^{l,lf,la} r k d>:
%% Premises:
%% <(e0^l0 e1^l1)^{l,lf,la} r k d> in conf.cs
%% Thus:
%% <(e0^l0 e1^l1)^{l,lf,la} r s k d> -->lazy   <e0^l0 r s\sqcup[ld |-> {k}] ar^{l,lf,la}(e1^l1 r ld) d>
%% <(e0^l0 e1^l1)^{l,lf,la} r s k d> -->strict <e0^l0 r s\sqcup[ld |-> {k}] ar^{l,lf,la}(e1^l1 r ld) d>

%% Stores are same in this case.

%% <v ar^{l,lf,la}(e^le r a) d>:
%% Premises:
%% For all v' in force(v, s),
%%   <v' ar^{l,lf,la}(e^le r a) d> in conf.cs
%% Thus:
%% <v ar^{l,lf,la}(e^le r ka) d> -->lazy   <e^le r s\sqcup[lfd |-> force(v, s)] fn^{l,la}(lfd ka) d>
%% for all v' in force(v, s),
%% <v' ar^{l,lf,la}(e^le r ka) d> -->strict <e^le r s\sqcup[lfd |-> {v'}] fn^{l,la}(lfd ka) d>
%% Thus the union of all the strict stores will equal the lazy store.

%% <v fn^{l,la}(fa ka) d>:
%% Premises:
%% For all v' in force(v, s),
%%   <v' fn^{l,la}(fa ka) d> in conf.cs
%% Thus
%% <v fn^{l,la}(fa ka) d> -->lazy <e^fl r[x |-> xd'] s'\sqcup[xd' |-> s'(lad)] k d'>
%% where k in s(ka)
%%       clos(x e^fl r) in s(fa)
%%       d' = truncate(ld, K)
%%       s' = s\sqcup[lad |-> force(v, s)]
%% For all v' in force(v,s),
%% <v' fn^{l,la}(fa ka) d> -->strict <e^fl r[x |-> xd'] s'\sqcup[xd' |-> s'(lad)] k d'>
%% where k in s(ka)
%%       clos(x e^fl r) in s(fa)
%%       d' = truncate(ld, K)
%%       s' = s\sqcup[lad |-> {v'}]

%% Thus the union of all the strict stores will equal the lazy store.
